Measuring Perceptions and Preferences for Meritocracy

### 2020-09-30 docs/paper.Rmd
paper para proofreading

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
abstract: "Economic and social inequality have raised growing concerns and crises across societies. One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on criteria such as effort and talent. Despite its wide use, empirical research on meritocracy is something relatively novel. A number of studies have relied mostly in secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items. Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy. Based on this critical analysis, we propose an item's battery called *Perceptions and Preferences for Meritocracy Scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141). The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions. The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::pdf_document2:
    template: null
    toc: false
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
  - ../input/bib/EconomiaMoral.bib
  - ../input/bib/meritocracy.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

# Introducción

La desigualdad económica se ha vuelto un tema que genera creciente preocupación y malestar alrededor del mundo. Esto se ha expresado en una serie de protestas como la emblemática "occupy wall street" el año 2011, así como también en una serie de análisis críticos respecto del desarrollo del capitalismo y sus consecuencias [@piketty_Capital_2014]. En este contexto, el estudio de las visiones, preferencias y percepciones respecto de la desigualdad han adquirido relevancia en las ciencias sociales, en temas como las preferencias redistributivas [@alesina_Fairness_2005; @dimick_Models_2018] la legitimación de la desigualdad económica [@schroder_Income_2017]  y el funcionamiento de la meritocracia [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014].

En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962]. Esta concepción tradicional de mérito sitúa en un lugar secundario la posible interferencia de factores estructurales o no meritocráticos como la herencia, los contactos personales, o la suerte [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994]. Desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019]. Tales investigaciones han planteado críticas a la realización de este estándar moral de distribución, argumentando que sería una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019, @khan_Privilege_2013].

Debido al rol que cumplen las creencias meritocráticas dentro del pensamiento neoliberal [@bay-chengTrackingHomoOeconomicus2015], han surgido múltiples investigaciones que evalúan la relación entre creencias meritocráticas y diversos ámbitos sociales de la actualidad. Por ejemplo, se han desarrollado estudios que vinculan la meritocracia al reforzamiento de estereotipos socioeconómicos, de género y de etnias [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], así como también líneas de investigación que evalúan el efecto de las creencias meritocráticas en el contexto educativo [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] y en el contexto organizacional de las empresas [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].

Para poder dar cuenta de los niveles de creencia en la meritocracia los estudios a la fecha generalmente han utilizado algunos indicadores de encuestas ya existentes, y en menor medida se han creado instrumentos ad-hoc. Sin embargo, y como mostraremos más adelante, las formas de medición de meritocracia varían extremadamente entre estudios. Muchas veces fenómenos similares se asocian a indicadores distintos, y también ocurre que fenómenos distintos son medidos con indicadores similares, todo lo cual dificulta la comparabilidad entre investigaciones y avanzar en la comprensión y estudio de la meritocracia.

Basados en el análisis crítico de las formas de medición de meritocracia a la fecha, el presente artículo propone un instrumento para medir y relacionar dos aspectos claves en el estudio de la meritocracia: percepciones y preferencias. Además, como un segundo eje de análisis considera la generación de indicadores respecto de aspectos meritocráticos y anti-meritocráticos, demostrando que no son los dos polos de un mismo continuo como muchos estudios anteriores parecen sugerir. La propuesta de medición además está orientada a generar un instrumento lo más breve posible de manera que pueda ser utilizado en encuestas de opinión pública y así ser estudiado en relación a otros fenómenos sociales.

# La medición de los aspectos subjetivos de la meritocracia

A continuación se presenta una revisión de una serie de investigaciones que se han abocado al estudio de la meritocracia y que para ello han hecho una propuesta de medición. El primer eje de análisis tiene que ver con el uso del concepto "creencias" para referir a distintos aspectos subjetivos relacionados con meritocracia. El segundo eje tiene que ver con el uso de indicadores sobre aspectos anti-meritocráticos como el polo opuesto de los meritocráticos.

@@

Several approaches to the empirical study of meritocracy based on public opinion surveys make reference to the concept of _beliefs_, but behind this concept there are usually different meanings and operationalizations. To illustrate this point in the following we will start with the proposal from a recent paper by @mijs_paradox_2019, which we will take as a reference to discuss previous studies on meritocracy.

The meritocratic beliefs' definition of Mijs is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9]. In the operationalization, this is associated with the following indicator: "how important you think it is for getting ahead in life: (a) hard work", scored in a 1 to 5 likert scale. There are several assumptions behind this decision that are worth discussing and that are related to previous assumptions in the study of meritocracy.

a. _Dimensionality_

The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead". This battery presents a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, wealthy family, right connections, religion, race and gender. Therefore, for Mijs other aspects such as education, that could be associated to talent, are not meritocratic. As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation, ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is conditioned by a nonmeritocratic factor: who your parents happen to be" (p.5).

In Mijs' proposal we can observe a couple of strong assumptions: effort would not depend on parents influence, and talent is not meritocratic (contrary to Michael Youngs original conceptualization). The problem of whether talent is or not meritocratic is certainly an interesting point to discuss, but for those working based on empirical data it is something that should be empirically tested. Is effort the only dimension behind the concept of meritocracy, or is it a multidimensional concept that includes other elements as talent (as in its original conceptualization)? This conceptual and measurement unidimensional assumption it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].

@@

The "reasons to get ahead" battery refers to "how important you think it is", considered by @mijs_paradox_2019 as a belief in meritocracy for the item regarding effort. Nevertheless, another version of this same battery used in several surveys is "how important you think it _should_ be".  Therefore, the question raised here is: Which one of both is a "belief"? According to @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e. thoughts about what is). Beliefs are here defined as normative ideas about just inequality (i.e. thoughts about what should be)"(p.359). Therefore, the referred paper would be using the term beliefs (what should be) referring actually to perceptions (what is). The same occurs in studies as the one by @reynolds_perceptions_2014, who explicitly use the term beliefs to talk about perceptions, whereas other authors use general terms as attitudes (@kunovich_systems_2007). The first attempt to shed light on this confusion was made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy (beliefs), whereas for "perceived" meritocracy they use two items: “Would you say that in your country, people are rewarded for their efforts?” and “... people are rewarded for their skills?”.

Is the belief in meritocracy a perception or a desire/preference? In order to expand the analytical conceptual framework, we believe that both dimensions should be included in the analysis, as proposed by @duru-bellat_whos_2012. This opens up possibilities of analyzing whether perceptions and preferences are actually related  (i.e. correlation close to 1) or whether they are different aspects of the same phenomenon. As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p. 435). In other words, normative beliefs should be interpreted taken perception into account: a large normative belief in meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy. In order to avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences and meritocratic perceptions, as they better reflect the two dimension under scrutiny.

@@

@mijs_paradox_2019 makes reference to some non-meritocratic aspects as talent, which is ruled out of the operationalization of meritocracy. A different approach was followed by @kunovich_systems_2007, who decide to include some non-meritocratic elements. Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), they decided that reasons as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e. reverse coded). A similar approach was taken by @newman_false_2015, reverse-coding non-meritocratic items, the same principle applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].

The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the  "get ahead" perceptions' battery items mentioned above. They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections and non-mertitocratic. Nevertheless, despite making this distinction the author ends up subtracting one dimension from the other, assuming that they are two poles of the same continuum as @kunovich_systems_2007 did. Still, taking into account this research perspective, we suggest that non-meritocratic aspects should be part of a meritocratic measurement but taken independently and not adding or subtracting from meritocratic ones unless it is empirically proved that they belong to the same conceptual dimension.

@@

Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989], using single indicators and/or simple average indexes for measuring meritocracy. Such strategy assumes that the latent construct is measured perfectly by the indicators chosen, going as far as proposing that "... In choosing this strategy of index construction, we argue that *support for meritocracy is not a latent variable* [@kunovich_systems_2007 p.653-654]. Some advances were done by @reynolds_perceptions_2014 by doing a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily, they end up in a sum index despite proving a multidimensional latent structure.

@@

Based on the previous limitations and assumptions in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:

@@

The columns Perceptions and Preferences represent the distinction between this two concepts, usually confused under the label "beliefs". Perceptions refers to the extent to which people observe that meritocracy functions or applies in their society, which in terms of measurement relates to items such as "I think hard work is important to get ahead in society", whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g. whether hard work should be related to payment). The rows in the table of Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014]. This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requieres empirical scrutiny. These non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.

Regarding the selection of indicators, most of them are taken or adapted from previous studies for the sake of comparability. For meritocratic indicators we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962, whereas for non-meritocratic dimensions we use having rich parents and good contacts. The description of the specific items is presented in the methodology section.

@@

* $H_1$. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and the effort to get ahead in life.

@@

The data was obtained through an online questionnaire which was part of a larger study on meritocracy and preferences developed in Chile in 2019 and funded by the national scientific agency FONDECYT. The questionnaire was programmed in Qualtrics and the fieldwork was in charge of an external online survey agency ([netquest.cl](www.netquest.cl)) during December 2019 and January 2020. The sample was selected from a non-probabilistic quota design in three large cities in Chile. The quotas were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country and is also the counterpart agency of ISSP (International Social Survey Programme) in Chile. A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale and those who did not accept informed consent. As it usually occurs online samples, there were some limitations in achieving the quotas for lower educational levels.

@@

With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different order versions as explained in Figure \@ref(fig:appmod). The scale was presented to the first group (_n = 712_) in the order that appears in Table 2. For the second group (_n = 717_), the order of the items was organized according to the topics of the items, e.g. for the topic of hard work the item about perception was followed by the item about preference, and te same for the rest of the topics. Finally, for the third group (_n = 712_) the items were completely randomized.

@@

A pre-registration was made in the OSF platform, available in the following link: [https://osf.io/z45y2](https://osf.io/z45y2). In this pre-registration are the hypothesis regarding the four-dimensional conceptual model underlying the scale,  the variables measurement levels, the statistical tests to be performed with their respective evaluation parameters, along with other important aspects of the research design.

@@

Como puede observarse en la tabla \@ref(tab:desc01), los indicadores poseen valores que van desde el 1 (totalmente en desacuerdo) al 5 (totalmente de acuerdo). Se observan promedios desde 2.41 correspondiente a preferencia-contactos, hasta 3.89 correspondiente a la preferencia-esfuerzo. Ambos indicadores son coherentes con la adhesión general a la meritocracia reportada en estudios previos, privilegiando aspectos individuales como el esfuerzo.

```{r desc01, echo=FALSE}
cap <- "Descriptive statistics of the scale."
load(file = "../input/data/proc/dat04.RData")
dat04b<- dat04 %>% rename("A. Perception Effort" =perc_effort,
                          "B. Perception Talent" = perc_talent,
                          "C. Perception rich parents" = perc_wpart,
                          "D. Perception contacts" = perc_netw,
                          "E. Preferences Effort" = pref_effort,
                          "F. Preferences Talent" =  pref_talent,
                          "G. Preferences rich parents" =  pref_wpart ,
                          "H. Preferences contacts" = pref_netw  )
sum04b  <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = cap
) %>% kable_styling(full_width = table_format2,
                    font_size = 10,
                    position = "center",
                    latex_options = c("hold_position"));sumdat04
```

Los gráficos presentados en la Figura \@ref(fig:plotlikert) presentan información desagregada y comparable de las distintas categorías de respuesta de cada item.  Se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto. En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento como criterio de preferencia meritocrático.

@@

En la Figura \@ref(fig:corpoly) se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia por esfuerzo y por talento, r=0.56). Destacan además las relaciones entre las percepciones y preferencias de tipo meritocráticas, lo que no ocurre con los indicadores no meritocraticos.

@@

\@ref(fig:meas02) SHOWS the results of the estimation for the four-factor model with complete sample. This model presents adequate fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88. Furthermore, considering the correlation between the latent variables, it is observed that the variables with the highest correlations are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not present a considerable correlation.

>>>>

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
abstract: "Economic and social inequality have raised growing concerns and crises across contemporary societies. One of the mechanisms developed by the social sciences to explain the maintenance of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences of effort and talent. Despite its wide use, empirical research on meritocracy is relatively novel and since it is based on multiple approaches and techniques its analysis can be vague or confusing. Particularly problematic is that a number of studies have relied mostly on secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items. Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy. Based on this critical analysis, we propose an item's battery called *Perceptions and Preferences for Meritocracy Scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141). The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions. The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::pdf_document2:
    template: null
    toc: false
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
  - ../input/bib/EconomiaMoral.bib
  - ../input/bib/meritocracy.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

# Introduction

Economic inequality has become an issue of growing concern around the world. This has been expressed in a series of protests, as well as diverse critical analyses regarding the development of capitalism and its consequences [Streeck, 2014; @piketty_Capital_2014]. In this context, the study of visions, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005]; Dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; Atria et al., 2020].

In general, meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a sum of talent and effort [@young_rise_1962]. This traditional conception of merit places in a secondary position the possible interference of structural or nonmeritocratic factors such as inheritance, personal contacts, or luck [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994]. Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, in general with the hypothesis that greater belief in meritocracy leads to greater legitimization of inequalities [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019]. Such research has raised criticism of the realization of this moral standard of distribution, arguing that it would be an unfulfilled promise given the preponderant influence of other elements beyond merit on individual status [Sandel, 2020; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019, @khan_Privilege_2013].

Due to the role that meritocratic beliefs play as a justification for individual achievement or failure in contemporary societies [@bay-chengTrackingHomoOeconomicus2015], multiple investigations have emerged that evaluate the relationship between meritocratic beliefs and diverse social circumstances today. For example, studies have been developed that link meritocracy to the reinforcement of socio-economic, gender and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; premingerMeritocracyServiceEthnocracy2020], as well as lines of research that evaluate the effect of meritocratic beliefs in the educational context [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and in the organizational dimension of companies [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].

To account for the levels of belief in meritocracy, studies have generally used some indicators from existing surveys, and to a lesser extent have created ad-hoc instruments. However, as we will show later, the forms of measuring meritocracy vary extremely among studies. Many times similar phenomena are associated with different indicators, and it also happens that different phenomena are measured with similar indicators, all of which make difficult not only the comparability between studies but also to acheive more precision in understanding the effects of meritocratic beliefs in different fields.

Based on the critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a tool to measure and relate two key aspects in the study of meritocracy: perceptions and preferences. Furthermore, as a second axis of analysis it considers the elaboration of indicators regarding meritocratic and non-meritocratic aspects, demonstrating that they are not the two poles of the same continuum as many previous studies seem to suggest. The measurement proposal is also oriented to generate an instrument as brief as possible so that it can be used in public opinion surveys, integrating meritocratic beliefs in the study of different social phenomena.

@@

Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but behind this concept there are usually different meanings and operationalizations. To illustrate this point in the following we will take as a main reference a recent paper about meritocracy by @mijs_paradox_2019, which makes several assumptions about the measurement of meritocracy that are also present in previous studies.

The meritocratic beliefs' definition of Mijs is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9]. In the operationalization, this is associated with the following indicator: "how important you think it is for getting ahead in life: (a) hard work", scored in a 1 to 5 likert scale. There are several assumptions behind this decision that are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.

a. _Unidimensionality_

The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead". This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, wealthy family, right connections, religion, race and gender. Therefore, for Mijs other aspects such as education, that could be associated to talent, are not meritocratic. As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation, ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is conditioned by a nonmeritocratic factor: who your parents happen to be" (p.5).

In this measure of meritocratic belief we can observe a couple of strong assumptions: effort would not depend on parents influence, and talent is not meritocratic (contrary to Michael Youngs original conceptualization). The problem of whether talent is or not meritocratic is certainly an interesting point to discuss, but for those working based on empirical data it is something that should be empirically tested. Is effort the only dimension behind the concept of meritocracy, or is it a multidimensional concept that includes other elements as talent (as in its original conceptualization)? This conceptual and measurement unidimensional assumption it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019]. Recently, Mijs & Hoy (2020) considered a measurement of talent as an indicator of meritocracy.

@@

The most used item battery for the operationalization of meritocracy using survey data is the "reasons to get ahead", which asks about "how important you think it is" and then lists several reasons such as effort, education, parents wealth and contacts. Nevertheless, another version of this same battery used in several surveys is "how important you think it _should_ be".  Therefore, the question raised here is: Which one of both is a "belief"?

The term belief has an ambiguous character in the literature, conceived as "idea-elements" for Converse (1964) or "considerations" for Zaller (1992). As Kluegel & Smith pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30). Therefore, it covers almos anything related to subjective factors. But a relevant distinction in this field was made by @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e. thoughts about what is). Beliefs are here defined as normative ideas about just inequality (i.e. thoughts about what should be)"(p.359). Several papers dealing with meritocracy use the term beliefs (what should be) referring actually to perceptions (what is). This happens not only in mijs_paradox_2019, but also in @reynolds_perceptions_2014, who explicitly use the term beliefs to talk about perceptions, whereas other authors use general terms as attitudes (@kunovich_systems_2007). The first attempt to shed light on this issue was made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy (beliefs), whereas for "perceived" meritocracy they used two items: “Would you say that in your country, people are rewarded for their efforts?” and “... people are rewarded for their skills?”.

Is the belief in meritocracy a perception or a preference with normative meaning? In order to expand the analytical conceptual framework, we believe that both dimensions should be included in the analysis, as proposed by @duru-bellat_whos_2012. This opens up possibilities of analyzing whether perceptions and preferences are actually related  (i.e. correlation close to 1) or whether they are different aspects of the same phenomenon. As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p. 435). In other words, normative beliefs should be interpreted taking perceptions into account: a large normative belief in meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy. Avoiding the confusion generated by the term "belief", we propose the terms meritocratic preferences (normative dimension) and meritocratic perceptions (factual dimention), as they better reflect the two faces of meritocracy under scrutiny.

@@

Unlike studies that consider talent a non-meritocratic aspect and therefore rule it out of the operationalization of meritocracy, @kunovich_systems_2007 follow a different approach by including some non-meritocratic elements. Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), they argue that reasons such as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e. reverse coded). A similar approach was taken by @newman_false_2015, reverse-coding non-meritocratic items, the same principle applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].

The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the  "get ahead" perceptions' battery items mentioned above. They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections as non-meritocratic. Nevertheless, despite making this distinction the author ends up subtracting one dimension from the other, assuming that they are two poles of the same continuum as @kunovich_systems_2007 did. Still, taking into account this view, we suggest that non-meritocratic aspects should be part of a meritocratic measurement but taken independently and not adding or subtracting from meritocratic ones unless it is empirically proved that they belong to the same conceptual dimension.

@@

Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J. M., 2008), using single indicators and/or simple average indexes for measuring meritocracy. Such strategy assumes that the latent construct is measured perfectly by the indicators chosen, going as far as proposing that "... In choosing this strategy of index construction, we argue that *support for meritocracy is not a latent variable* [@kunovich_systems_2007 p.653-654]. Some advances were done by @reynolds_perceptions_2014 by doing a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily, they end up in a sum index despite proving a multidimensional latent structure.

@@

Based on the previous assumptions and limitations in the measurement of meritocracy, we propose and test an instrument with the following characteristics:

@@

The columns Perceptions and Preferences represent the distinction between these two concepts, usually confused under the label "beliefs". Perceptions refers to the extent to which people observe that meritocracy works or applies in their society, which in terms of measurement relates to items such as "I think hard work is important to get ahead in society", whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g. whether hard work should be related to payment). The rows in the table of Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014]. This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requires empirical scrutiny. These non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.

Regarding the selection of indicators, most of them are taken or adapted from previous studies for the sake of comparability. For meritocratic indicators we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962 (Atria et al. 2020) , whereas for non-meritocratic dimensions we use having rich parents and good contacts. The description of the specific items is presented in the methodology section.

@@

* $H_1$. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and effort to get ahead in life.

@@

The data was obtained through an online questionnaire which was part of a larger study on meritocracy and preferences developed in Chile in 2019 and funded by the national scientific agency (ANID). The questionnaire was programmed in Qualtrics and the fieldwork was in charge of an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020. The sample was selected from a non-probabilistic quota design in three large cities in Chile. The quotas for gender, age and education level were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country and is also the counterpart agency of ISSP (International Social Survey Programme) in Chile. A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale and those who did not accept informed consent. There are not significant differences between our effective sample and the population for most socio-demographics characteristics, with the exception of educational level (see Table XX in appendix). As it usually occurs with online samples, there were some limitations in achieving the quotas for lower educational levels.

@@

With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different set-order versions as explained in Figure \@ref(fig:appmod). The scale was presented to the first group (_n = 712_) in the order that appears in Table 2. For the second group (_n = 717_), the order of the items was organized according to the topics of the items, e.g. for the topic of hard work the item about perception was followed by the item about preference, and the same for the rest of the topics. Finally, for the third group (_n = 712_) the items were completely randomized.

@@

A pre-registration was made in the OSF platform, available in the following link: [https://osf.io/z45y2](https://osf.io/z45y2). In this pre-registration are the hypothesis regarding the four-dimensional conceptual model underlying the scale, the variables measurement levels, and the statistical tests to be performed with their respective evaluation parameters along with other important aspects of the research design.

@@

As can be seen in the table \ref(tab:desc01), the indicators have values ranging from 1 (totally disagree) to 5 (totally agree). Averages are observed from 2.41, corresponding to preference-contacts, to 3.89, corresponding to preference-effort. Both indicators are consistent with the general adherence to meritocracy reported in previous studies, privileging individual aspects such as effort (Frei et al 2020, Bucca, 2016).

```{r desc01, echo=FALSE}
cap <- "Descriptive statistics of the scale."
load(file = "../input/data/proc/dat04.RData")
dat04b<- dat04 %>% rename("A. Perception Effort" =perc_effort,
 "B. Perception Talent" = perc_talent,
 "C. Perception rich parents" = perc_wpart,
 "D. Perception contacts" = perc_netw,
 "E. Preferences Effort" = pref_effort,
 "F. Preferences Talent" = pref_talent,
 "G. Preferences rich parents" = pref_wpart ,
 "H. Preferences contacts" = pref_netw )
sum04b <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
sumdat04<- kable(table04b,format =table_format, booktabs= T, exhaust = FALSE,align = "l",linesep = "", caption = cap
) %>% kable_styling(full_width = table_format2,
 font_size = 10,
 position = "center",
 latex_options = c("hold_position"));sumdat04
```

The graphs presented in Figure \@ref(fig:plotlikert) present disaggregated and comparable information of the different response categories for each item. It can be seen that in general there is more agreement in the perception of non meritocratic items than meritocratic ones, while in the case of preferences the opposite occurs. As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.

@@

In Figure Ref(fig:corpoly), moderate to high intensity relationships are observed between indicators that correspond to the same factor (for example, perception of meritocracy by effort and by talent, r=0.56). The relationships between meritocratic perceptions and preferences are also noteworthy, which is not the case for nonmeritocratic indicators.

@@

\@ref(fig:meas02) shows the results of the estimation for the four-factor model with complete sample. This model presents adequate fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88. Furthermore, considering the correlation between the latent variables, it is observed that the variables with the highest correlations are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not present a considerable correlation.
